Proof-Reconstruction in Propositional Logic

### 2017-10-20 paper/sections/proof-reconstruction.tex
[ paper ] function names using \rm.

Therefore, for a general recursive function $f: \Prop \to \Prop$
with termination problems, we state a new function
$f^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must be
done by using induction on the \Nat type structure.

@@

\begin{align}
  \label{eq:unshunt-definition}
  \begin{split}
    &unshunt_{0} : \Prop → \Prop\\
    &unshunt_{0}(φ) =\\
    &\hspace{3mm}\begin{cases}
    unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
    unshunt_{0}(φ₁ ⇒ φ₂) ∧ unshunt_{0}(φ₁ ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
    φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
\end{align}

@@

Thus, we define the function
$unshunt_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:unshunt-structured}).

\begin{align}
  \label{eq:unshunt-structured}
  \begin{split}
    &unshunt_{1} : \Prop → \Nat \to \Prop\\
    &unshunt_{1}(φ, \zero)\hspace{3mm}= φ\\%\hspace{10cm}\\
    &unshunt_{1}(φ, \suc(n)) =\\
    &\hspace{3mm}\begin{cases}
    unshunt_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
      &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
    unshunt_{1}(φ₁ ⇒ φ₂, n)\\
    ∧\ unshunt_{1}(φ₁ ⇒ φ₃, n),
       &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
  \end{cases}
  \end{split}
\end{align}

As we mentioned above, the natural number in the second argument of
the function used in Eq.~(\ref{eq:unshunt-structured}) stands for a
complexity measure of the function, in this case $unshunt_{1}$ function.

Recall the \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number for a proposition $φ$ like the following $rank$ function.
The symbols $c_{¬}$ and $c_{\square}$ are non-negative integer
numbers called the complexity of the respective connective.

\begin{align*}
  \begin{split}
  % \label{eq:rank-definition}
    &rank : \Prop → \Nat\\
    &rank(φ) =
    \begin{cases}
    0, &\text{for atomic }φ \\
    rank(φ₁) + c_{¬},  &\text{ if } φ ≡ \neg φ₁ \\
    \max{(rank(φ₁),\ rank(φ₂))} + c_{\square},
      &\text{ if } φ ≡ φ₁\,\square\, φ₂.
    \end{cases}
  \end{split}
\end{align*}

We define the $unshunt_{cm}: \Prop → \NN$ function as
the complexity measure for the $unshunt_0$ function.

\begin{align*}
  \label{eq:unshunt-rank-definition}
  \begin{split}
      &unshunt_{cm} : \Prop → \Nat\\
      &unshunt_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        unshunt_{cm}(φ₃) + 2, &\text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \max(unshunt_{cm}(φ₂),\ unshunt_{cm}(φ₃)) + 1,
          &\text{ if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{ otherwise.}
        \end{cases}
  \end{split}
\end{align*}

Finally, following the technique mentioned, we get a curated version of
$unshunt_{0}$ function presented in Def.~\ref{def:unshunt}.

\begin{definition}[unshunt]
  \label{def:unshunt}
    \begin{align*}
      \begin{split}
      &unshunt : \Prop → \Prop\\
      &unshunt(\varphi) = unshunt_{1}(\varphi, unshunt_{cm}(\varphi)).
      \end{split}
    \end{align*}
\end{definition}

@@

\begin{itemize}
\item \texttt{rule} is a $n$-ary \Metis inference rule
\item $rule$ is the emulated version function  for
the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &rule : \Prop → \Prop → \cdots → \Prop\\
    &rule(φ₁,\ \cdots, φₙ, ψ) =
      \begin{cases}
      ψ, &\text{ if }rule\text{ built }ψ\text{ by applying valid}\\
         &\text{ transformations or other rules to }φ\\
      φ, &\text{ otherwise.}
      \end{cases}
  \end{split}
\end{align*}

@@

@@

\begin{lemma}[lem-unshunt]
  \label{lem:unshunt}
  $Γ ⊢ unshunt(φ)$, then $Γ ⊢ φ$.
\end{lemma}

\begin{sketchproof} Use induction on the cases defined by the outcome of the
$unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
			  argument.
\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n) ∧ unshunt(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-unshunt,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ φ₂) ∧ unshunt(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-unshunt.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

Finally, using the theorem \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{sketchproof}

We now define the $split$ function that yields the conjunction of subgoals
that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[split]
\label{def:split}
  \begin{equation*}
    \begin{split}
    &split : \Prop → \Prop\\
    &\mysplit(φ) =
      \begin{cases}
      unshunt(\mysplit(φ₁)) &\text{ if }φ ≡ φ₁ ∧ φ₂\\
      ∧\ unshunt(φ₁ ⇒ \mysplit(φ₂)), &\\[2.5mm]
      unshunt(¬ φ₁ ⇒ \mysplit(φ₂)),
        & \text{ if }φ ≡ φ₁ ∨ φ₂\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(φ₂)),
        & \text{ if }φ ≡ φ₁ ⇒ φ₂\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(φ₂))
        & \text{ if }φ ≡ φ₁ ⇔ φ₂\\
      ∧\ unshunt(φ₂ ⇒ \mysplit(φ₁)),
        &\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∧ φ₂)\\[2.5mm]
      unshunt(\mysplit(¬ φ₁))
        & \text{ if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
      ∧\ unshunt(¬ φ₁ ⇒ \mysplit(¬ φ₂)),
        &\\[2.5mm]
      unshunt(\mysplit(φ₁))
        & \text{ if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
      ∧\ unshunt(φ₁ ⇒ \mysplit(¬ φ₂)),
        &\\[2.5mm]
      unshunt(φ₁ ⇒ \mysplit(¬ φ₂))
        & \text{ if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
      ∧\ unshunt(¬ φ₂ ⇒ \mysplit(φ₁)),
        &\\[2.5mm]
      unshunt(\mysplit(φ₁)),
        & \text{ if }φ ≡ ¬ ¬ φ₁\\
      ⊤,
        & \text{ if }φ ≡ ¬ ⊥\\
      ⊥,
        & \text{ if }φ ≡ ¬ ⊤.
      \end{cases}
    \end{split}
  \end{equation*}
\end{definition}

The definition above has termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such an issue following the technique
described in Section~\ref{ssec:structural-recursion}. We refer to $split$
as the function after applying the process mentioned above here and below
unless otherwise stated.

\begin{lemma}[\thmsplit] $Γ ⊢ \mysplit(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-split}
\end{lemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula $φ$ by following the cases in Def.~\ref{def:split}.
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.
\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(split(¬¬φ₁,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢split(¬¬φ₁,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁)$}
  \RightLabel{lem-unshunt.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁) ∧ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ⇒ φ₂)$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{axiom}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

@@

  \AxiomC{$Γ ⊢ \mysplit(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ unshunt(¬ φ₁ ⇒ \mysplit(φ₂,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \mysplit(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \mysplit(φ₂,n-1)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

  \AxiomC{$\mysplit(φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \mysplit(φ)$}
  \RightLabel{\thmsplit.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

First, we define the $rdisj$ function to convert
a disjunction in its right-associative form.
The $rconj$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

>>>>

Therefore, for a general recursive function $\rm{f}: \Prop \to \Prop$
with termination problems, we state a new function
$\rm{f}^{*} : \Prop \to \Nat \to \Prop$ where all recursive calls must be
done by using induction on the \Nat type structure.

@@

\begin{align}
  \label{eq:uh-definition}
  \begin{split}
    &\fuh_{0} : \Prop → \Prop\\
    &\fuh_{0}(φ) =\\
    &\hspace{3mm}\begin{cases}
    \fuh_{0}((φ₁ ∧ φ₂) ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
    \fuh_{0}(φ₁ ⇒ φ₂) ∧ \fuh_{0}(φ₁ ⇒ φ₃),
      & \text{ if }φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
    φ, &\text{ otherwise.}
    \end{cases}
  \end{split}
\end{align}

@@

Thus, we define the function
$uh_{1}$ by structural recursion in its second argument in
Eq.~(\ref{eq:uh-structured}).

\begin{align}
  \label{eq:uh-structured}
  \begin{split}
    &\fuh_{1} : \Prop → \Nat \to \Prop\\
    &\fuh_{1}(φ, \zero)\hspace{3mm}= φ\\%\hspace{10cm}\\
    &\fuh_{1}(φ, \suc(n)) =\\
    &\hspace{3mm}\begin{cases}
    \fuh_{1}((φ₁ ∧ φ₂) ⇒ φ₃, n),
      &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\[2mm]
    \fuh_{1}(φ₁ ⇒ φ₂, n)\\
    ∧\ \fuh_{1}(φ₁ ⇒ φ₃, n),
       &\text{ if } φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃).
  \end{cases}
  \end{split}
\end{align}

As we mentioned above, the natural number in the second argument of
the function used in Eq.~(\ref{eq:uh-structured}) stands for a
complexity measure of the function, in this case $uh_{1}$ function.

Recall the \emph{complexity measure} defined in \cite{Agudelo-Agudelo2017}
and presented in \cite{VanDalen1994} as \emph{the rank for a
proposition}, is a function $\Prop \to \Nat$ that assigns a natural
number for a proposition $φ$ like the following $\frank$ function.
The symbols $c_{¬}$ and $c_{\square}$ are non-negative integer
numbers called the complexity of the respective connective.

\begin{align*}
  \begin{split}
  % \label{eq:rank-definition}
    &\frank : \Prop → \Nat\\
    &\frank(φ) =
    \begin{cases}
    0, &\text{for atomic }φ \\
    \frank(φ₁) + c_{¬},  &\text{ if } φ ≡ \neg φ₁ \\
    \fmax{(\frank(φ₁),\ \frank(φ₂))} + c_{\square},
      &\text{ if } φ ≡ φ₁\,\square\, φ₂.
    \end{cases}
  \end{split}
\end{align*}

We define the $\fuh_{cm}: \Prop → \NN$ function as
the complexity measure for the $\fuh_0$ function.

\begin{align*}
  \label{eq:uh-rank-definition}
  \begin{split}
      &\fuh_{cm} : \Prop → \Nat\\
      &\fuh_{cm}(φ) =\\
      &\hspace{3mm}
        \begin{cases}
        \fuh_{cm}(φ₃) + 2, &\text{ if }φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃)\\
        \fmax(\fuh_{cm}(φ₂),\ \fuh_{cm}(φ₃)) + 1,
          &\text{ if }  φ ≡ φ₁ ⇒ (φ₂ ∧ φ₃)\\
        0, &\text{ otherwise.}
        \end{cases}
  \end{split}
\end{align*}

Finally, following the technique mentioned, we get a curated version of
$\fuh_{0}$ function presented in Def.~\ref{def:uh}.

\begin{definition}[uh]
  \label{def:uh}
    \begin{align*}
      \begin{split}
      &\fuh : \Prop → \Prop\\
      &\fuh(φ) = \fuh_{1}(φ, \fuh_{cm}(φ)).
      \end{split}
    \end{align*}
\end{definition}

@@

\begin{itemize}
\item \texttt{rule} is a $n$-ary \Metis inference rule
\item $\rm{rule}$ is the emulated version function  for
the \texttt{rule} inference.
\begin{align*}
  \begin{split}
    &\rm{rule} : \Prop → \Prop → \cdots → \Prop\\
    &\rm{rule}(φ₁,\ \cdots, φₙ, ψ) =
      \begin{cases}
      ψ, &\text{ if }\rm{rule}\text{ built }ψ\text{ by applying valid}\\
         &\text{ transformations or other rules to }φ\\
      φ, &\text{ otherwise.}
      \end{cases}
  \end{split}
\end{align*}

@@

% ------------------------------------------------------------------------------

@@

\begin{lemma}[lem-inv-uh]
  \label{lem:uh}
  $Γ ⊢ \fuh(φ)$, then $Γ ⊢ φ$.
\end{lemma}

\begin{sketchproof} Use induction on the cases defined by the outcome of the
$\fuh$ function in Eq.~(\ref{eq:uh-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
			  argument.
\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒\rm{-to-}⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

\item $(φ₁ ⇒ (φ₂ ∧ φ₃))$
\begin{equation*}
  \scalebox{0.95}{
  ($\mathcal{D}_1$)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n) ∧ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{∧-proj₁}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂, n-1)$}
    \RightLabel{lem-inv-uh,}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
  }
\end{equation*}

\begin{equation*}
  (\mathcal{D}_2)
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \fuh(φ₁ ⇒ φ₂) ∧ \fuh(φ₁ ⇒ φ₃)$}
    \RightLabel{∧-proj₂}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n)$}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ φ₃, n-1)$}
    \RightLabel{lem-inv-uh.}
    \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₃$}
    \end{bprooftree}
\end{equation*}

Finally, using the theorem \texttt{⇒∧⇒\rm{-to-}⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒\rm{-to-}⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒\rm{-to-}⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{sketchproof}

We now define the $\fstrip$ function that yields the conjunction of
subgoals that implies the goal of the problem in the \Metis \TSTP derivations.

\begin{definition}[strip]
\label{def:strip}
  \begin{equation*}
    \begin{split}
    &\fstrip : \Prop → \Prop\\
    &\fstrip(φ) =
      \begin{cases}
      \fuh(\fstrip(φ₁)) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂)),
      &\text{ if }φ ≡ φ₁ ∧ φ₂ \\
      \fuh(¬ φ₁ ⇒ \fstrip(φ₂)),
        & \text{ if }φ ≡ φ₁ ∨ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(φ₂)),
        & \text{ if }φ ≡ φ₁ ⇒ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(φ₂)) ∧ \fuh(φ₂ ⇒ \fstrip(φ₁)),
        & \text{ if }φ ≡ φ₁ ⇔ φ₂\\
      \fuh(φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∧ φ₂)\\
      \fuh(\fstrip(¬ φ₁)) ∧\ \fuh(¬ φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ∨ φ₂)\\
      \fuh(\fstrip(φ₁)) ∧\ \fuh(φ₁ ⇒ \fstrip(¬ φ₂)),
        & \text{ if }φ ≡ ¬ (φ₁ ⇒ φ₂)\\
      \fuh(φ₁ ⇒ \fstrip(¬ φ₂)) ∧\ \fuh(¬ φ₂ ⇒ \fstrip(φ₁)),
        & \text{ if }φ ≡ ¬ (φ₁ ⇔ φ₂)\\
      \fuh(\fstrip(φ₁)),
        & \text{ if }φ ≡ ¬ ¬ φ₁\\
      ⊤,
        & \text{ if }φ ≡ ¬ ⊥\\
      ⊥,
        & \text{ if }φ ≡ ¬ ⊤.\\
      φ, &\text{ otherwise}.
      \end{cases}
    \end{split}
  \end{equation*}
\end{definition}

The definition above has termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such an issue following the technique
described in Section~\ref{ssec:structural-recursion}.
We refer to $\fstrip$ as the function after applying the process mentioned above here and below unless otherwise stated.

\begin{lemma}[lem-inv-strip] $Γ ⊢ \fstrip(φ)$ then $Γ ⊢ φ$.
\label{lem:lem-inv-strip}
\end{lemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula $φ$ by following the cases in Def.~\ref{def:strip}.
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.

\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(\fstrip(¬¬φ₁,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ \fstrip(¬¬φ₁,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁)$}
  \RightLabel{lem-inv-uh.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁) ∧ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ⇒ φ₂)$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{}
  \RightLabel{assume φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{weaken φ₁}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ , φ₁ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip}
  \UnaryInfC{$Γ , φ₁ ⊢ φ₂$}
  \RightLabel{⇒-intro.}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip(φ, n)$}
\RightLabel{axiom}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh(¬ φ₁)$}
\RightLabel{lem-inv-uh,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

@@

  \AxiomC{$Γ ⊢ \fstrip(φ,n-1)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁) ∧ \fuh(¬ φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh(¬ φ₁ ⇒ \fstrip(φ₂,n-1))$}
  \RightLabel{lem-inv-uh}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip(φ₂,n-1)$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip(φ₂,n-1)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

  \AxiomC{$\fstrip(φ) ≡ \bigwedge_{i=1}^n s_i $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip(φ)$}
  \RightLabel{lem-inv-strip.}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

@@

First, we define the $\rm{rdisj}$ function to convert
a disjunction in its right-associative form.
The $\rm{rconj}$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).
